How do changes in ex vary between countries?

### 2022-03-01 draft_paper.Rmd
looked at changes in ex by decade

---
title: "How do changes in ex vary between countries?"
author: "Jon Minton"
date: "22/02/2022"
output:
  word_document: default
  html_document: default
---

@@

change_in_ex_selected_countries
```

```{r}
change_in_ex_selected_countries %>% 
  filter(x == 0) %>% 
  left_join(country_code_lookup) %>% 
  mutate(country = factor(country, levels = c("England & Wales", "Scotland", "East Germany", "West Germany", "Germany", "Synthetic Germany", "Spain", "France", "Italy"))) %>% 
  filter(between(year, 1980, 2020)) %>% 
  ggplot(aes(x = year, y = delta_ex)) + 
  geom_point(alpha = 0.2) + 
  stat_smooth() + 
  facet_grid(country~sex) + 
  geom_hline(yintercept = 0) + 
  theme(strip.text.y = element_text(angle = 0)) + 
  labs(
    x = "Year",
    y = "Year-on-year change in life expectancy",
    title = "Annual change in life expectancy at birth, selected countries",
    subtitle = "Blue line: nonlinear smoother",
    caption =  "Source: Human Mortality Database. Synthetic Germany based on 20% East/80% West German population weighting"
  )

@@

>>>>

---
title: "How do changes in ex vary between countries?"
author: "Jon Minton"
date: "22/02/2022"
output:
  html_document: default
  word_document: default
---

@@

change_in_ex_selected_countries
```
### Changes in life expectancy at birth

```{r, echo = FALSE, warning=FALSE, message = FALSE, fig.height=8, fig.width=8}
change_in_ex_selected_countries %>% 
  filter(x == 0) %>% 
  left_join(country_code_lookup) %>% 
  mutate(country = factor(country, levels = c("England & Wales", "Scotland", "East Germany", "West Germany", "Germany", "Synthetic Germany", "Spain", "France", "Italy"))) %>% 
  filter(between(year, 1980, 2020)) %>% 
  ggplot(aes(x = year, y = delta_ex)) + 
  geom_point(alpha = 0.2) + 
  stat_smooth() + 
  facet_grid(country~sex) + 
  geom_hline(yintercept = 0) + 
  theme(strip.text.y = element_text(angle = 0)) + 
  labs(
    x = "Year",
    y = "Year-on-year change in life expectancy",
    title = "Annual change in life expectancy at birth, selected countries",
    subtitle = "Blue line: nonlinear smoother",
    caption =  "Source: Human Mortality Database. Synthetic Germany based on 20% East/80% West German population weighting"
  )

### Changes in life expectancy at age 65

```{r, echo = FALSE, warning=FALSE, message = FALSE, fig.height=8, fig.width=8}
change_in_ex_selected_countries %>% 
  filter(x == 65) %>% 
  left_join(country_code_lookup) %>% 
  mutate(country = factor(country, levels = c("England & Wales", "Scotland", "East Germany", "West Germany", "Germany", "Synthetic Germany", "Spain", "France", "Italy"))) %>% 
  filter(between(year, 1980, 2020)) %>% 
  ggplot(aes(x = year, y = delta_ex, colour=sex)) + 
  geom_point(alpha = 0.2) + 
  stat_smooth() + 
  facet_grid(country~sex) + 
  geom_hline(yintercept = 0) + 
  theme(strip.text.y = element_text(angle = 0)) + 
  labs(
    x = "Year",
    y = "Year-on-year change in life expectancy",
    title = "Annual change in life expectancy at age 65, selected countries",
    subtitle = "Blue line: nonlinear smoother",
    caption =  "Source: Human Mortality Database. Synthetic Germany based on 20% East/80% West German population weighting"
  )

@@

# Time trends

```{r}

hmd_ex_selected_countries_with_synth %>% 
    filter(year >= 1979) %>% 
    group_by(code, x, sex) %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
    nest() %>% 
    mutate(
      rwd_model = map(data, ~lm(delta_ex ~ 1, data = . )),
      ar_model  = map(data, ~lm(delta_ex ~ lag(delta_ex), data = .))
    ) %>% 
    mutate(
      aic_rwd = map_dbl(rwd_model, AIC),
      aic_ar  = map_dbl(ar_model, AIC)
    ) %>% 
    mutate(
      diff_aic = aic_ar - aic_rwd
    )  %>% 
  # filter(diff_aic > 0)
  mutate(
    tidied_ar_model = map(ar_model, broom::tidy)
  ) %>% 
  unnest(tidied_ar_model) %>% 
  filter(term == "(Intercept)") %>% 
  arrange(desc(estimate)) %>% 
  select(code, x, sex, estimate, std.error) %>% 
  mutate(display_label = glue::glue("{code}_{x}_{sex}")) %>% 
  ggplot(aes(estimate, fct_reorder(display_label, estimate))) + 
  geom_point()

Look at improvements by decade

```{r}

hmd_ex_selected_countries_with_synth %>% 
    filter(year >= 1979) %>% 
    group_by(code, x, sex) %>% 
    arrange(year) %>% 
    mutate(decade = year - (year %% 10)) %>% 
    mutate(delta_ex = ex - lag(ex)) %>%
    nest() %>%
    mutate(
      rwd_model = map(data, ~lm(delta_ex ~ factor(decade), data = . ))
     )  %>%
  mutate(
    tidied_rwd_model = map(rwd_model, broom::tidy)
  ) %>%
  unnest(tidied_rwd_model)
```
